---
title: "Report_Old_Data"
author: "Yunyao Zhu"
date: "4/15/2021"
output: pdf_old_document
number_sections: true
fig_caption: yes
header-includes: 
  - \usepackage{breqn}
  - \usepackage{amsmath}
bibliography: sources.bib
nocite: '@*'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
library(tidyverse)
library(lubridate) # extract month
library(forecast)
library(car)
library(tseries) # adf.test (Augmented Dickey-Fuller Test)
```

## Exploratory Data Analysis

```{r}
df_old = read.csv('avocado.csv')
# https://stackoverflow.com/questions/4310326/convert-character-to-date-in-r
# https://stackoverflow.com/questions/22603847/how-to-extract-month-from-date-in-r

df_old$Date = as.Date(df_old$Date, "%Y-%m-%d")
df_old$Month = month(df_old$Date)
df_old$Year.Month <- format(as.Date(df_old$Date), "%Y-%m")
df_old = df_old[order(df_old$Date), ]

df_old = df_old %>%
  rename(
    Average.Price = 'AveragePrice',
    Small = 'X4046',
    Large = 'X4225',
    XLarge = 'X4770',
    Type = 'type',
    Year = 'year',
    Region = 'region'
  )
```

```{r}
# https://stackoverflow.com/questions/4310326/convert-character-to-date-in-r
# https://stackoverflow.com/questions/22603847/how-to-extract-month-from-date-in-r
df_old$Date = as.Date(df_old$Date, "%Y-%m-%d")
df_old$Month = month(df_old$Date)
df_old$Year.Month <- format(as.Date(df_old$Date), "%Y-%m")
df_old = df_old[order(df_old$Date), ]
```

```{r}
df_old.total.weekly.conventional = df_old[(df_old$Region == 'TotalUS') & (df_old$Type == 'conventional'), ]
df_old.total.weekly.organic = df_old[(df_old$Region == 'TotalUS') & (df_old$Type == 'organic'), ]
```

```{r}
df_old.total.monthly.conventional = df_old.total.weekly.conventional %>%
                              group_by(Year.Month) %>%
                              summarise_at(vars(Date, Average.Price, Total.Volume, Small, Large, XLarge, Total.Bags, Small.Bags, Large.Bags, XLarge.Bags), funs(mean(., na.rm=TRUE)))


df_old.total.monthly.organic = df_old.total.weekly.organic %>%
                              group_by(Year.Month) %>%
                              summarise_at(vars(Date, Average.Price, Total.Volume, Small, Large, XLarge, Total.Bags, Small.Bags, Large.Bags, XLarge.Bags), funs(mean(., na.rm=TRUE)))
```

```{r}
df_old.total.weekly.conventional$Total.Volume.Million = df_old.total.weekly.conventional$Total.Volume/1000000
df_old.total.weekly.organic$Total.Volume.Million = df_old.total.weekly.organic$Total.Volume/1000000
df_old.total.monthly.conventional$Total.Volume.Million = df_old.total.monthly.conventional$Total.Volume/1000000
df_old.total.monthly.organic$Total.Volume.Million = df_old.total.monthly.organic$Total.Volume/1000000
```













```{r}
ggtsdisplay(old_monthly_conventional_volume, points = FALSE, lag.max = 36)
```

```{r}
ggtsdisplay(diff(df_old.total.monthly.conventional$Total.Volume.Million, difference = 1), points = FALSE, lag.max = 36)
```
```{r}
ggtsdisplay(diff(df_old.total.monthly.conventional$Total.Volume.Million, difference = 2), points = FALSE, lag.max = 36)
```
```{r}
ggtsdisplay(old_monthly_conventional_price, points = FALSE, lag.max = 36)
```

```{r}
ggtsdisplay(diff(df_old.total.monthly.conventional$Average.Price, difference = 1), points = FALSE, lag.max = 36)
```
```{r}
ggtsdisplay(diff(df_old.total.monthly.conventional$Average.Price, difference = 2), points = FALSE, lag.max = 36)
```



# EDA

## Weekly Volume
```{r fig.height = 2.8, fig.width = 6, fig.cap="Scatter Plot of Per-Unit Price v.s. Total Volume of Avocados Sold in Millions"}
ggplot(data = df_old.total.weekly.conventional, aes(x = Average.Price, y = Total.Volume.Million)) +
  geom_point() +
  geom_smooth(method = "lm", col='grey') + 
  labs(title = 'Potential Inverse Relationship between Total Volume and Price',
       x = 'Per-Unit Price ($)',
       y = 'Total Number of Avocados Sold (in Millions)')
```

```{r}
m0 = lm(Total.Volume.Million ~ Average.Price, data = df_old.total.weekly.conventional)
summary(m0)
plot(m0)
```

```{r}
plot(fitted.values(m0), residuals(m0), xlab = "Fitted Values", ylab = "Residuals", 
main = "Heteroscedasticity in Residuals vs. Fitted Value plot")
abline(h=0, col="red")
```

```{r}
hist(residuals(m0), xlab = "Residuals", ylab = "Frequency", main = "Residuals Slightly Positively Skewed")
```

```{r fig.height = 2.8, fig.width = 6, fig.cap="Scatter Plot of Per-Unit Price v.s. Total Volume of Avocados Sold"}
ggplot(data = df_old.total.weekly.conventional, aes(x = Year.Month, y = Total.Volume.Million)) +
  geom_point() +
  geom_smooth(method = "lm", col='grey') + 
  labs(title = 'Potential Inverse Relationship between Total Volume and Price',
       x = 'Per-Unit Price ($)',
       y = 'Total Number of Avocados Sold')
```

```{r fig.height = 5, fig.width = 7, fig.cap="Time Series Plot of Total Volume of Conventional Avocados in the U.S."}
# time series plots: 
# https://otexts.com/fpp2/ts-objects.html
# https://otexts.com/fpp2/stl.html
# https://otexts.com/fpp2/classical-decomposition.html

old_weekly_conventional_volume <- ts(df_old.total.weekly.conventional$Total.Volume.Million, start = c(2015, 1), frequency = 52)

old_weekly_conventional_volume %>% decompose(type="multiplicative") %>%
  autoplot() + xlab("Year") +
  ggtitle("Volume of Conventional Avocados Sold in the U.S. Show Seasonal Pattern")
```

```{r fig.height = 3.5, fig.width = 7.2, fig.cap="Predicting the Weekly Per-Unit Conventional Avocado Prices from 2017-2018"}
man.fit.week <- Arima(old_weekly_conventional_volume[1:117], xreg=(df_old.total.weekly.conventional[,"Average.Price"])[1:117], order = c(2,0,0), seasonal = list(order = c(0,1,1), period = 52))
man.pre.week <- predict(man.fit.week, newxreg=(df_old.total.weekly.conventional[,"Average.Price"])[118:169],  n.ahead=52)
```

```{r}
# par(cex=0.7, mfrow=c(1,2)) 
# par(mai=c(0.8,0.5,0.8,0.2)) # https://datavoreconsulting.com/post/spacing-of-panel-figures-in-r/
matplot(1:52, cbind(old_weekly_conventional_volume[118:169], man.pre.week$pred), type = "l", xlab = "Weeks (03/2017-03/2018)\n(Prediction Period Only)", ylab = "Predicted Volume (in Millions)", main = "One-Year Sales Volume Prediction")
abline(v=c(10, 20, 35, 48), col=c("grey", "grey", "grey", "grey"), lty=c(3, 3, 3, 3), lwd=c(1, 1, 1, 1))
```


```{r}
# http://www.sthda.com/english/wiki/abline-r-function-an-easy-way-to-add-straight-lines-to-a-plot-using-r-software
par(mai=c(0.8,0.5,0.8,0.2))
fmanweek1 <- forecast(man.fit.week, xreg=(df_old.total.weekly.conventional[,"Average.Price"])[118:169], h=52, level = c(95))
{plot(fmanweek1, main = "One-Year Prediction with 95% CI", xlab = 'Weeks (03/2017-03/2018)\n(Total Time Span)')
lines(ts(old_weekly_conventional_volume))}
```

## Monthly Volume
```{r fig.height = 2.8, fig.width = 6, fig.cap="Scatter Plot of Per-Unit Price v.s. Total Volume of Avocados Sold in Millions"}
ggplot(data = df_old.total.monthly.conventional, aes(x = Average.Price, y = Total.Volume.Million)) +
  geom_point() +
  geom_smooth(method = "lm", col='grey') + 
  labs(title = 'Potential Inverse Relationship between Total Volume and Price',
       x = 'Per-Unit Price ($)',
       y = 'Total Number of Avocados Sold (in Millions)')
```

```{r}
m0 = lm(Total.Volume.Million ~ Average.Price, data = df_old.total.monthly.conventional)
summary(m0)
plot(m0)
```

```{r}
plot(fitted.values(m0), residuals(m0), xlab = "Fitted Values", ylab = "Residuals", 
main = "Heteroscedasticity in Residuals vs. Fitted Value plot")
abline(h=0, col="red")
```
```{r}
residuals(m0)
```


```{r}
plot(residuals(m0), xlab = "Month", ylab = "Residuals", 
main = "Heteroscedasticity in Residuals vs. Fitted Value plot")
# abline(h=0, col="red")
```

```{r}
hist(residuals(m0), xlab = "Residuals", ylab = "Frequency", main = "Residuals Slightly Positively Skewed")
```

```{r}
m1 = lm(Total.Volume.Million ~ Average.Price + Earnings + Unemployment + CPI, data = df_old.total.monthly.conventional.merged)
summary(m1)
plot(m1)
```
```{r}
vif(m2)
```

```{r}
m2 = lm(Total.Volume.Million ~ Average.Price + Unemployment + CPI, data = df_old.total.monthly.conventional.merged)
summary(m2)
plot(m2)
```


```{r}
plot(fitted.values(m2), residuals(m2), xlab = "Fitted Values", ylab = "Residuals", 
main = "Heteroscedasticity in Residuals vs. Fitted Value plot")
abline(h=0, col="red")
```

```{r}
plot(residuals(m2), xlab = "Fitted Values", ylab = "Residuals", 
main = "Heteroscedasticity in Residuals vs. Fitted Value plot")
abline(h=0, col="red")
```

```{r}
hist(residuals(m2), xlab = "Residuals", ylab = "Frequency", main = "Residuals Slightly Positively Skewed")
```
```{r fig.height = 5, fig.width = 7, fig.cap="Time Series Plot of Total Volume of Conventional Avocados in the U.S."}
# time series plots: 
# https://otexts.com/fpp2/ts-objects.html
# https://otexts.com/fpp2/stl.html
# https://otexts.com/fpp2/classical-decomposition.html

m2_residuals <- ts(residuals(m2), frequency = 12)

m2_residuals %>% decompose() %>%
  autoplot() + xlab("Year") +
  ggtitle("Residuals Show Seasonal Pattern")
```

```{r}
adf.test(m2_residuals, alternative = "stationary")
```


```{r}
ggtsdisplay(m2_residuals, points = FALSE, lag.max = 39)
adf.test(m2_residuals, alternative = "stationary")
```

```{r}
ggtsdisplay(diff(m2_residuals, difference = 1), points = FALSE, lag.max = 39)
adf.test(diff(m2_residuals, difference = 1), alternative = "stationary")
```
```{r}
ggtsdisplay(diff(m2_residuals, difference = 2), points = FALSE, lag.max = 39)
adf.test(diff(m2_residuals, difference = 2), alternative = "stationary")
```

```{r}
ggtsdisplay(diff(m2_residuals, difference = 3), points = FALSE, lag.max = 39)
adf.test(diff(m2_residuals, difference = 3), alternative = "stationary")
```













```{r fig.height = 2.8, fig.width = 6, fig.cap="Scatter Plot of Per-Unit Price v.s. Total Volume of Avocados Sold"}
ggplot(data = df_old.total.monthly.conventional, aes(x = Year.Month, y = Total.Volume.Million)) +
  geom_point() +
  geom_smooth(method = "lm", col='grey') + 
  labs(title = 'Potential Inverse Relationship between Total Volume and Price',
       x = 'Per-Unit Price ($)',
       y = 'Total Number of Avocados Sold')
```

```{r fig.height = 5, fig.width = 7, fig.cap="Time Series Plot of Total Volume of Conventional Avocados in the U.S."}
# time series plots: 
# https://otexts.com/fpp2/ts-objects.html
# https://otexts.com/fpp2/stl.html
# https://otexts.com/fpp2/classical-decomposition.html

old_monthly_conventional_volume <- ts(df_old.total.monthly.conventional$Total.Volume.Million, start = c(2015, 1), frequency = 12)

old_monthly_conventional_volume %>% decompose(type="multiplicative") %>%
  autoplot() + xlab("Year") +
  ggtitle("Volume of Conventional Avocados Sold in the U.S. Show Seasonal Pattern")
```

```{r}
adf.test(old_monthly_conventional_volume, alternative = "stationary")
```


```{r}
adf.test(old_weekly_conventional_volume, alternative = "stationary")

```

```{r}
week_price.diff <- diff(avocado_price$AveragePrice, difference = 1)
adf.test(week_price.diff,alternative = "stationary")
```


```{r}
df_old.total.monthly.conventional$Total.Volume.Million %>%
  ggtsdisplay(xlab="Month",
    main="Total Volume")
```


```{r}
(fr_m1 = Arima(old_monthly_conventional_volume, xreg=df_old.total.monthly.conventional[["Average.Price"]], order = c(0,0,0),
seasonal = list(order=c(0,0,0), period=12)))
# man.fit.month <- Arima(old_monthly_conventional_volume[1:27], xreg=df_old.total.monthly.conventional[["Average.Price"]][1:27], order = c(2,0,0), seasonal = list(order = c(1,1,1), period = 12))

```


```{r}
(fr_m2 = Arima(old_monthly_conventional_volume, xreg=df_old.total.monthly.conventional[["Average.Price"]], order = c(0,0,0),
seasonal = list(order=c(0,1,0), period=12)))
```

```{r}
(fr_m3.1 = Arima(old_monthly_conventional_volume, xreg=df_old.total.monthly.conventional[["Average.Price"]], order = c(0,0,0),
seasonal = list(order=c(0,1,1), period=12)))
```

```{r}
(fr_m3.2 = Arima(old_monthly_conventional_volume, xreg=df_old.total.monthly.conventional[["Average.Price"]], order = c(0,0,0),
seasonal = list(order=c(0,1,2), period=12)))
```


```{r}
(fr_m3.3 = Arima(old_monthly_conventional_volume, xreg=df_old.total.monthly.conventional[["Average.Price"]], order = c(0,0,0),
seasonal = list(order=c(0,1,3), period=12)))
```

```{r}
(fr_m4.1 = Arima(old_monthly_conventional_volume, xreg=df_old.total.monthly.conventional[["Average.Price"]], order = c(1,0,0),
seasonal = list(order=c(0,1,2), period=12)))
```

```{r}
(fr_m4.2 = Arima(old_monthly_conventional_volume, xreg=df_old.total.monthly.conventional[["Average.Price"]], order = c(2,0,0),
seasonal = list(order=c(0,1,2), period=12)))
```


```{r}
summary(fr_m4.2)
```


```{r}
df_old.total.monthly.conventional$Total.Volume.Million %>% diff(lag=1) %>%
  ggtsdisplay(xlab="Month",
    main="Seasonally differenced Total Volume")
```


```{r}
## ACF - Seasonal AR (Q)
acf(df_old.total.monthly.conventional$Total.Volume.Million, lag.max = 12,main = "")
```

[Rephrase]: From the correlogram we can see that there is significant correlation (blue line represents 95% confidence) from lag 1, therefore an AR order (Q) of 1 is trialled.


```{r}
# Differenced ACF - Non-Seasonal AR (q)
old_monthly_conventional_volume.diff <- diff(df_old.total.monthly.conventional$Total.Volume.Million, difference = 1)
acf(old_monthly_conventional_volume.diff, lag.max = 12,main = "")
```
[Rephrase:] As only lag 0 is significant this could suggest that AR of the non-seasonal component (q) is 0.

```{r}
## PACF - Seasonal MA (P)
pacf(df_old.total.monthly.conventional$Total.Volume.Million, lag.max = 12,main = "")
```

Again, here we see a correlation at only lag 0/2, therefore a MA (P) of 0/2 is used in the seasonal component.???

```{r}
pacf(old_monthly_conventional_volume.diff, lag.max=12,main = "")
```

Finally, using the difference PACF, we can work out the AR (p) of the non-seasonal order. It shows that only the first 3 lags are significantly correlated, therefore 3 is used.


Seasonal ARIMA $(p, d, q) \times (P, D, Q)_{s}$


```{r fig.height = 3.5, fig.width = 7.2, fig.cap="Predicting the Weekly Per-Unit Conventional Avocado Prices from 2017-2018"}
man.fit.month <- Arima(old_monthly_conventional_volume[1:27], xreg=df_old.total.monthly.conventional[["Average.Price"]][1:27], order = c(2,0,0), seasonal = list(order = c(0,1,2), period = 12))
man.pre.month <- predict(man.fit.month, newxreg=df_old.total.monthly.conventional[["Average.Price"]][28:39],  n.ahead=12)
```

```{r}
# par(cex=0.7, mfrow=c(1,2)) 
# par(mai=c(0.8,0.5,0.8,0.2)) # https://datavoreconsulting.com/post/spacing-of-panel-figures-in-r/
matplot(1:12, cbind(old_monthly_conventional_volume[28:39], man.pre.month$pred), type = "l", xlab = "Months (03/2017-03/2018)\n(Prediction Period Only)", ylab = "Predicted Volume (in Millions)", main = "One-Year Sales Volume Prediction")
# abline(v=c(10, 20, 35, 48), col=c("grey", "grey", "grey", "grey"), lty=c(3, 3, 3, 3), lwd=c(1, 1, 1, 1))
```


```{r}
# par(cex=0.7, mfrow=c(1,2)) 
# par(mai=c(0.8,0.5,0.8,0.2)) # https://datavoreconsulting.com/post/spacing-of-panel-figures-in-r/
matplot(1:12, cbind(old_monthly_conventional_volume[28:39], man.pre.month$pred, old_monthly_conventional_volume[15:27]), type = "l", xlab = "Months (03/2017-03/2018)\n(Prediction Period Only)", ylab = "Predicted Volume (in Millions)", main = "One-Year Sales Volume Prediction")
# abline(v=c(10, 20, 35, 48), col=c("grey", "grey", "grey", "grey"), lty=c(3, 3, 3, 3), lwd=c(1, 1, 1, 1))
```

```{r}
fmanweek1
```



```{r}
# http://www.sthda.com/english/wiki/abline-r-function-an-easy-way-to-add-straight-lines-to-a-plot-using-r-software
# par(mai=c(0.8,0.5,0.8,0.2))
fmanweek1 <- forecast(man.fit.month, xreg=df_old.total.monthly.conventional[["Average.Price"]][28:39], h=12, level = c(95))
{plot(fmanweek1, main = "Prediction with 95% CI", xlab = 'Months (01/2015-03/2018)\n(Total Time Span)', ylab = "Predicted Volume (in Millions)")
lines(ts(old_monthly_conventional_volume))}
```

```{r}
summary(man.fit.month)
```

```{r}
forecast::ggtsdisplay(fr_m4.2$residuals[28:39], points=FALSE, lag.max = 36)

```

```{r}
checkresiduals(man.fit.month)
```









```{r fig.height = 3.5, fig.width = 7.2, fig.cap="Predicting the Weekly Per-Unit Conventional Avocado Prices from 2017-2018"}
man.fit.month <- Arima(old_monthly_conventional_volume[1:27], order = c(2,0,0), seasonal = list(order = c(0,1,2), period = 12))
man.pre.month <- predict(man.fit.month, n.ahead=12)
```

```{r}
# par(cex=0.7, mfrow=c(1,2)) 
# par(mai=c(0.8,0.5,0.8,0.2)) # https://datavoreconsulting.com/post/spacing-of-panel-figures-in-r/
matplot(1:12, cbind(old_monthly_conventional_volume[28:39], man.pre.month$pred), type = "l", xlab = "Months (03/2017-03/2018)\n(Prediction Period Only)", ylab = "Predicted Volume (in Millions)", main = "One-Year Sales Volume Prediction")
# abline(v=c(10, 20, 35, 48), col=c("grey", "grey", "grey", "grey"), lty=c(3, 3, 3, 3), lwd=c(1, 1, 1, 1))
```


```{r}
# http://www.sthda.com/english/wiki/abline-r-function-an-easy-way-to-add-straight-lines-to-a-plot-using-r-software
# par(mai=c(0.8,0.5,0.8,0.2))
fmanweek1 <- forecast(man.fit.month, h=12, level = c(95))
{plot(fmanweek1, main = "One-Year Prediction with 95% CI", xlab = 'Months (03/2017-03/2018)\n(Total Time Span)')
lines(ts(old_monthly_conventional_volume))}
```








## Weekly Price

```{r fig.height = 5, fig.width = 7, fig.cap="Time Series Plot of Per-Unit Price of Conventional Avocados in the U.S."}
# time series plots: 
# https://otexts.com/fpp2/ts-objects.html
# https://otexts.com/fpp2/stl.html
# https://otexts.com/fpp2/classical-decomposition.html

old_weekly_conventional_price <- ts(df_old.total.weekly.conventional$Average.Price, start = c(2015, 1), frequency = 52)

old_weekly_conventional_price %>% decompose(type="multiplicative") %>%
  autoplot() + xlab("Year") +
  ggtitle("Per-Unit Price of Conventional Avocados in the U.S. Show Seasonal Pattern")
```


## Monthly Price
```{r fig.height = 5, fig.width = 7, fig.cap="Time Series Plot of Per-Unit Price of Conventional Avocados in the U.S."}
# time series plots: 
# https://otexts.com/fpp2/ts-objects.html
# https://otexts.com/fpp2/stl.html
# https://otexts.com/fpp2/classical-decomposition.html

old_monthly_conventional_price <- ts(df_old.total.monthly.conventional$Average.Price, start = c(2015, 1), frequency = 12)

old_monthly_conventional_price %>% decompose(type="multiplicative") %>%
  autoplot() + xlab("Year") +
  ggtitle("Per-Unit Prices of Conventional Avocados in the U.S. Show Seasonal Pattern")
```

```{r}
df.earnings.month = read.csv('average_weekly_earnings.csv')[ ,c('Year', 'Label', 'Value')]
df.earnings.month$Month = substr(df.earnings.month$Label, 6, 8)
df.earnings.month$Month = match(df.earnings.month$Month, month.abb)
df.earnings.month = df.earnings.month[ ,c('Year', 'Month', 'Value')]

df.earnings.month$Year.Month = paste(as.character(df.earnings.month$Year), sprintf("%02d",df.earnings.month$Month), sep="-")

df.earnings.month = df.earnings.month %>% rename(Earnings = 'Value')
```


```{r}
# https://stackoverflow.com/questions/21982987/mean-per-group-in-a-data-frame
df.unemployment.month = read.csv('unemployment_rate.csv')[ ,c('Year', 'Label', 'Value')]
df.unemployment.month$Month = substr(df.unemployment.month$Label, 6, 8)
df.unemployment.month$Month = match(df.unemployment.month$Month, month.abb)
df.unemployment.month = df.unemployment.month[ ,c('Year', 'Month', 'Value')]

df.unemployment.month$Year.Month = paste(as.character(df.unemployment.month$Year), sprintf("%02d",df.unemployment.month$Month), sep="-")

df.unemployment.month = df.unemployment.month %>% rename(Unemployment = 'Value')
```

```{r}
df.cpi.month = read.csv('consumer_price_index_all_items.csv')
df.cpi.month$Date = as.Date(df.cpi.month$DATE, "%m/%d/%y")
df.cpi.month$Year = year(df.cpi.month$Date)
df.cpi.month$Month = month(df.cpi.month$Date)
df.cpi.month = df.cpi.month[ ,c('Year', 'Month', 'USACPIALLMINMEI', 'Date')]

df.cpi.month$Year.Month = paste(as.character(df.cpi.month$Year), sprintf("%02d",df.cpi.month$Month), sep="-")

df.cpi.month = df.cpi.month %>% rename(CPI = 'USACPIALLMINMEI')
```

```{r}
df_old.total.monthly.conventional.merged = cbind(df_old.total.monthly.conventional, df.earnings.month['Earnings'], df.unemployment.month['Unemployment'], df.cpi.month['CPI'])
```


```{r fig.height = 3, fig.width = 6, fig.cap="Scatter Plot of Per-Unit Price v.s. Total Volume of Avocados Sold"}
ggplot(data = df_old.total.monthly.conventional.merged, aes(x = Earnings, y = Average.Price)) +
  geom_point() +
  geom_smooth(method = "lm", col='grey') + 
  labs(title = 'Potential Correlation between Per-Unit Prices and Earnings',
       x = 'Average Weekly Earnings ($)',
       y = 'Per-Unit Prices ($)')
```


```{r fig.height = 3, fig.width = 6, fig.cap="Scatter Plot of Per-Unit Price v.s. Total Volume of Avocados Sold"}
ggplot(data = df_old.total.monthly.conventional.merged, aes(x = Unemployment, y = Average.Price)) +
  geom_point() +
  geom_smooth(method = "lm", col='grey') + 
  labs(title = 'Potential Inverse Relationship between Per-Unit Prices and Unemployment Rates',
       x = 'Unemployment Rates (%)',
       y = 'Per-Unit Prices ($)')
```


```{r fig.height = 3, fig.width = 6, fig.cap="Scatter Plot of Per-Unit Price v.s. Total Volume of Avocados Sold"}
ggplot(data = df_old.total.monthly.conventional.merged, aes(x = CPI, y = Average.Price)) +
  geom_point() +
  geom_smooth(method = "lm", col='grey') + 
  labs(title = 'Potential Correlation Between CPI and Unemployment Rates',
       x = 'Consumer Price Indices (CPI)',
       y = 'Per-Unit Prices ($)')
```


```{r}
m4 = lm(Average.Price ~ Total.Volume.Million + Earnings + Unemployment + CPI, data = df_old.total.monthly.conventional.merged)
summary(m4)
plot(m4)
```
```{r}
vif(m4)
```


```{r}
m5 = lm(Average.Price ~ Total.Volume.Million + Unemployment + CPI, data = df_old.total.monthly.conventional.merged)
summary(m5)
plot(m5)
```

```{r}
vif(m5)
```



```{r}
covariates.mat = cbind(df_old.total.monthly.conventional.merged[['Earnings']],
             df_old.total.monthly.conventional.merged[['Unemployment']],
             df_old.total.monthly.conventional.merged[['CPI']])
```



```{r}
adf.test(old_monthly_conventional_price, alternative = "stationary")
```

```{r}
old_monthly_conventional_price.diff <- diff(df_old.total.monthly.conventional$Average.Price, difference = 2)
adf.test(old_monthly_conventional_price.diff,alternative = "stationary")
```


```{r}
df_old.total.monthly.conventional$Average.Price %>%
  ggtsdisplay(xlab="Month",
    main="Total Volume")
```
```{r}
(fr_m0.0 = Arima(old_monthly_conventional_price, xreg=covariates.mat, order = c(0,0,0),
seasonal = list(order=c(0,0,0), period=12)))
# Arima(old_monthly_conventional_price[1:27], xreg=covariates.mat[1:27]
# man.fit.month <- Arima(old_monthly_conventional_volume[1:27], xreg=df_old.total.monthly.conventional[["Average.Price"]][1:27], order = c(2,0,0), seasonal = list(order = c(1,1,1), period = 12))

```


```{r}
(fr_m0.1 = Arima(old_monthly_conventional_price, xreg=covariates.mat, order = c(0,1,0),
seasonal = list(order=c(0,0,0), period=12)))
```


```{r}
(fr_m0.2 = Arima(old_monthly_conventional_price, xreg=covariates.mat, order = c(0,2,0),
seasonal = list(order=c(0,0,0), period=12)))
```


```{r}
(fr_m1 = Arima(old_monthly_conventional_price, xreg=covariates.mat, order = c(0,1,0),
seasonal = list(order=c(0,0,0), period=12)))
# man.fit.month <- Arima(old_monthly_conventional_volume[1:27], xreg=df_old.total.monthly.conventional[["Average.Price"]][1:27], order = c(2,0,0), seasonal = list(order = c(1,1,1), period = 12))

```


```{r}
(fr_m2 = Arima(old_monthly_conventional_price, xreg=covariates.mat, order = c(0,1,0),
seasonal = list(order=c(0,1,0), period=12)))
```


```{r}
(fr_m3.1 = Arima(old_monthly_conventional_price, xreg=covariates.mat, order = c(0,1,0),
seasonal = list(order=c(0,0,1), period=12)))
```

```{r}
(fr_m3.2 = Arima(old_monthly_conventional_price, xreg=covariates.mat, order = c(0,1,0),
seasonal = list(order=c(0,0,2), period=12)))
```


```{r}
(fr_m3.3 = Arima(old_monthly_conventional_price, xreg=covariates.mat, order = c(0,1,0),
seasonal = list(order=c(0,0,3), period=12)))
```

```{r}
(fr_m4.1 = Arima(old_monthly_conventional_price, xreg=covariates.mat, order = c(1,0,0),
seasonal = list(order=c(0,1,2), period=12)))
```

```{r}
(fr_m4.2 = Arima(old_monthly_conventional_price, xreg=covariates.mat, order = c(2,0,0),
seasonal = list(order=c(0,1,2), period=12)))
```





```{r}
df_old.total.monthly.conventional$Average.Price %>% diff(lag=1) %>%
  ggtsdisplay(xlab="Month",
    main="Seasonally differenced Average Price")
```


```{r}
## ACF - Seasonal AR (Q)
acf(df_old.total.monthly.conventional$Average.Price, lag.max = 12,main = "")
```

[Rephrase]: From the correlogram we can see that there is significant correlation (blue line represents 95% confidence) from lag 1, therefore an AR order (Q) of 1 is trialled.


```{r}
# Differenced ACF - Non-Seasonal AR (q)
old_monthly_conventional_price.diff <- diff(df_old.total.monthly.conventional$Average.Price, difference = 1)
acf(old_monthly_conventional_price.diff, lag.max = 12,main = "")
```
[Rephrase:] As only lag 0 is significant this could suggest that AR of the non-seasonal component (q) is 0.

```{r}
## PACF - Seasonal MA (P)
pacf(df_old.total.monthly.conventional$Average.Price, lag.max = 12,main = "")
```

Again, here we see a correlation at only lag 0, therefore a MA (P) of 0 is used in the seasonal component.

```{r}
pacf(old_monthly_conventional_price.diff, lag.max=12,main = "")
```

Finally, using the difference PACF, we can work out the AR (p) of the non-seasonal order. It shows that only the first 3 lags are significantly correlated, therefore 3 is used.











```{r fig.height = 3, fig.width = 7.2, fig.cap="Predicting the Monthly Per-Unit Conventional Avocado Prices from 2017-2018"}
man.fit.month.price <- Arima(old_monthly_conventional_price[1:27], xreg=covariates.mat[1:27], order = c(1,1,0), seasonal = list(order = c(0,1,2), period = 12))
man.pre.month.price <- predict(man.fit.month.price, newxreg=covariates.mat[28:39], n.ahead=12)
```

```{r}
# par(cex=0.7, mfrow=c(1,2))
# par(mai=c(0.8,0.5,0.8,0.2)) # https://datavoreconsulting.com/post/spacing-of-panel-figures-in-r/
matplot(1:12, cbind(old_monthly_conventional_price[28:39], man.pre.month.price$pred), type = "l", xlab = "Months (03/2017-03/2018)\n(Prediction Period Only)", ylab = "Predicted Per-Unit Price ($)", main = "One-Year Per-Unit Price Prediction")
# abline(v=c(3, 11), col=c("grey", "grey"), lty=c(3, 3), lwd=c(1, 1))
# http://www.sthda.com/english/wiki/abline-r-function-an-easy-way-to-add-straight-lines-to-a-plot-using-r-software
```

```{r}
# par(mai=c(0.8,0.5,0.8,0.2))
fmanmonth1 <- forecast(man.fit.month.price, xreg=covariates.mat[28:39], h=12, level = c(95))
{plot(fmanmonth1, main = "One-Year Prediction of Per-Unit Price with 95% CI", xlab = 'Months (01/2015-03/2018)\n(Total Time Span)')
lines(ts(old_monthly_conventional_price))}
```



